Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-12 ms.tex
editing

\begin{abstract}
Steady-state one-dimensional hydrodynamic profiles of hot gas in the nuclei of quiescent galaxies are calculated for a range of black hole masses, parameterized gas heating rates, and physically-motivated stellar density profiles.  Mass is supplied to the circumnuclear medium (CNM) by stellar winds, while gas is heated by a combination of stellar winds, supernovae, and black hole feedback.  Analytic estimates are derived, neglecting radiative cooling, for quantities including the stagnation radius where the radial velocity goes to zero (if it exists) and the black hole accretion rate as a function of the heating efficiency, the latter of which is related to the star-formation history of the stellar population.  These results are used to assess over what range of conditions radiative instability will develop, assuming a feedback due to the average star formation history for galaxies of the given black hole mass.  Our main conclusion is that.....  Implications of our results for estimates of the accretion rate using Bondi formula, and for the diversity of CNM environments encountered by the jetted outflows from tidal disruption events, and for the use of an assumed $\dot{M}-M_{\bullet}$ relationship to constrain the occupation fraction of SMBHs in low mass galaxies, are briefly discussed.
\end{abstract}

@@

To understand why most SMBHs appear to be inactive requires understanding their gaseous environments.  Hot gas near the
SMBH sphere of influence (hereafter the `circumnuclear medium', or CNM) controls the central mass accretion rate $\dot{M}$.  The latter in turn affects the luminosity of the accretion flow and the possible impact of its energy and momentum input (`feedback') on larger scales.  The presence of dense gas in the nucleus can also lead to runaway cooling instabilities, resulting in bursty star-formation or episodic AGN activity (reference Ciotti and others).

Knowledge of how $\dot{M}$ varies with the SMBH mass $\Mbh$ and other properties of the
galactic nucleus informs key questions related to co-evolution of SMBHs and their host galaxies with cosmic time.  SMBH growth in the low redshift universe appears to be dominated by lower mass black holes (e.g.~\citealt{Heckman+04}).  This is often
attributed to cosmological `down sizing' (e.g, \citealt{Gallo+08}), even though the physical processes involved may be distinct from those operating at higher SMBH masses and accretion rates in quasars.

The accretion rates of low mass SMBHs, and their resulting nuclear X-ray luminosities $L_X$, also have implications for observational
constraints on their occupation fraction in low mass galaxies and the
possible existence of intermediate mass black holes.  By combining the average relationship between $\Mbh$ and
$L_X$ measured from a sample of high mass galaxies (\citealt{Gallo+08}) with upper limits on $L_X$ in lower mass
galaxies, \citet{MillerGallo+:2014a}  infer tentative evidence for an occupation fraction less than unity in the latter.  This method, however, relies on extrapolating a power-law fit to the $L_X-\Mbh$ distribution to low $\Mbh$ below the detection threshold.

@@

Past studies, including multi-dimensional hydrodynamics and a variety of (parameterized) physical effects, have focused on relatively massive elliptical galaxies (Ciotti refs).  A key result of this work is the periodic development of cooling instabilities, which temporarily increase the gas accretion rate before black hole feedback grows sufficiently important to halt this growth.  Here we instead focus on time-independent models, neglecting radiative cooling.  This simple approach allows us to systematically and to derive simple analytic expressions that prove useful in exploring whether cooling instabilties are always manifest, or whether other forms of non-AGN feedback can result in steady accretion.  Even if cooling instabilties develop on galactic scales, phases of stable accretion may persist on small scales.  We are thus motivated to explore whether it steady-flow is possible under any circumstances, or whether accretion is always building towards instability.

In the presence of strong heating, one-dimensional steady-state flow is characterized by an inflow-outflow structure, with a critical radius known as the``stagnation radius" $\rs$ where the radial velocity passes through zero.  Mass loss from stars interior to the stagnation radius is accreted, while that outside $\rs$ is unbound in an outflow from the nucleus.  The stagnation radius, rather than the Bondi radius, thus controls the SMBH accretion rate, although $\rs$ can in many cases lie within a factor of few of the nominal Bondi radius.  In the case of weak heating, however, no stagnation radius may exist at all, i.e. the entire ISM of the galaxy is feeding the SMBH.  Another goal of this work is to assess under what conditions a stagnation radius exists, as this question is intimately tied to the one of cooling instability.  
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion. 
The main goal of our analysis is to provide a framework for
interpreting the properties of quiescent galactic nuclei given the
comparatively well understood physics of stellar winds and supernova
heating.

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to measure the radial surface brightness profiles for hundreds of nearby early type galaxies. The measured brightness profile is generally well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.  Assuming the stellar population is spherically symmetric, this corresponds to a stellar density $\rhostar \propto r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for $r \gg \rb$.  To good approximation
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star(\rsoi) \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star(r_b) \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star(\rsoi)$ is the stellar density at the radius of the black hole sphere of influence (soi), 
\be
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc}.
\label{eq:rsoi}
\ee
The second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$ relationship of \citet{Gultekin+09},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
Although the use of equation (\ref{eq:Msigma}) is questionable for low mass black holes (e.g., \citealt{Greene&Ho07}), our results are not overly sensitive to this assumption.

A galaxy model is fully specified by four parameters: $\Mbh$, $\Gamma$, $r_b$,
and $\beta$.  We compute models for three different black hole masses, $\Mbh = 10^6$, $10^7$, $10^8 \Msun$.  The distribution of $\Gamma$ in the \citet{LauerFaber+:2007a} sample is bimodal, with a concentration of ``core" galaxies with $\Gamma \lsim 0.3$ and a concentration of ``cusp" galaxies with $\Gamma \gsim 0.7$.  We bracket these possibilities by considering models with $\Gamma=0.8$ and $\Gamma=0.1$.  Since we are generally interested in properties of the gas flow on radial scales $\ll r_b$, we fix $\beta = 2$ (a typical value), an assumption to which most of our results is not sensitive.  However, the presence of $\rb$ is necessary for to obtain a converged steady state for some regions of our parameter space. We adopt $\rb=100$ pc for cusp galaxies ($\Gamma=0.8$), while for core galaxy models we employ the scaling relation $\rb = $, motivated again by the \citet{LauerFaber+:2007a} sample. ({\bf BDM: fill in!}).  Though simplistic, our results are not sensitive to our assumption about the value of $\rb$.

@@

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the gas density profile $\rho$ of the CNM for each galaxy model by solving the equations of one-dimensional (spherically symmetric) time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial
      v}{\partial r}\right) =-\frac{\partial p}{\partial r}- \rho\frac{GM_{\rm enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] +\nabla\cdot(\kappa \nabla T)
, 
\label{eq:dsdt}
\end{align}
where $v$, $T$, $p$ and $s$ are the radial velocity, temperature,
pressure and specific entropy of the gas, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p =
\rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$.

The source term in the continuity equation (eq.~[\ref{eq:drhodt}]),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years.  To good approximation $\eta\simeq 0.02 (t_{\star}/t_h)^{-1.3}$ at time $t_{\star}$ following a burst of star formation (e.g., \citealt{Ciotti+91}),\footnote{\citet{Ciotti+91} give the overall mass return rate from evolved stars as a function of B-band luminosity.  Although we give the mass return rate per unit volume, our expressions are equivalent.} although $\eta$ is higher in cases of ongoing star formation.

Source terms $\propto q$ also appear in the momentum and entropy
equations (eqs.~[\ref{eq:dvdt}] and [\ref{eq:dsdt}]) because isotropic
mass injection in the SMBH rest frame represents a source of momentum
and energy relative to the mean flow.  Physically, these result from
the mismatch between the properties of virialized gas injected locally
by stellar winds and the mean properties of the background flow.  The
term $\propto -q c_{s}^{2}$ turns out to be an important in
stabilizing the flow to runaway cooling ($\S$).

The last term in equation (\ref{eq:dsdt}) accounts for heat
transportation by electron conductivity, where $\kappa_{\rm
  eff}=\kappa_{\rm spitz}/(1+\psi)$ (\citealt{DaltonBalbus:1993a}) is
the conducitivity and $\kappa_{\rm spitz} = \kappa_0 T^{5/2}$ is the
classical \citet{Spitzer62} value ($\kappa_0\simeq 2\times 10^{-6}$ in
cgs units).  The flux limitor $\psi = \kappa_{\rm spitz} \nabla
T/(5\phi \rho c_s^3)$ saturates the conductive flux if the mean free
path for electron coulomb scattering becomes comparable to the
temperature length scale, where $c_s \equiv (kT/\mu m_p)^{1/2}$ is the
isothermal sound speed and $\phi \lesssim 1$ is an uncertain
dimensionless constant (we adopt $\phi = 0.1$).  Even a weak magnetic
field perpendicular to the flow could in principle suppress the
conductivity.  However, the radially-decreasing temperature profiles
of interest are susceptible to the magneto-thermal instability
(\citealt{Balbus01}), the non-linear evolution of which is believed to
result in a radially-directed field structure
(\citealt{Parrish&Stone07}).

@@

@@

Radiative cooling is not included in our calculation.  Across much of
the parameter space radiative cooling is subdominant compared to other
sources of cooling.  Radiative cooling cannot be neglected when $\vwO$
is low (and/or $\eta$ is high), but once important it leads to thermal
instabilities (e.g.~\citealt{McCourt+12}) which our 1D
time-independent model cannot account for self-consistently anyways.
Indeed, one of the main goals of this work is to use our analytic and
numerical solutions neglecting radiation to ascertain over what
physical conditions cooling instabilities are likely to develop
($\S\ref{sec:cooling}$).

@@

\section{Results}

We begin by describing useful analytic results before presenting our numerical solutions.

\label{sec:results}
\subsection{Analytic Results}

Here we summarize analytic estimates of physical quantities, such as
the stagnation radius $\rs$ and the mass accretion rate, the detailed
derivation of which are given in Appendix \ref{app:rs}.  Our results
are expressed in terms of the SMBH mass $M_{\bullet} =
10^{8}M_{\bullet,8}M_{\odot}$ and the wind heating parameter
(eq.~[\ref{eq:dsdt}])
\begin{eqnarray}
  v_{\rm w} &=& v_{\rm w,0}\left(1+\frac{\sigma^{2}}{v_{\rm w,0}^{2}}\right)^{1/2}  \nonumber \\
  &\approx& v_{\rm w,0}\left(1+0.14 \Mbheight^{0.43}\left[\frac{v_{w,0}}{500\,\rm km\,s^{-1}}\right]^{-1}\right)^{1/2},
\end{eqnarray}
where we remind that the parameter $\vwO$ accounts for wind heating beyond the baseline level produced by the winds of old stars, and the final equality makes use of $M_{\bullet}-\sigma$ (eq.~[\ref{eq:Msigma}]).

\subsubsection{Flow Properties Near the Stagnation Radius}

Continuity of the entropy derivative at the stagnation radius demands that the temperature at the stagnation radius is given by (Appendix \ref{app:rs}, eq.~[\ref{eq:first_law}])
\begin{align}
T \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p v_{w}^{2}}{2} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K} 
\label{eq:Tanalytic}
\end{align}

Hydrostatic equilibrium can be used to determine the location of the stagnation radius where $v = 0$ (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
\rs=\frac{G \Mbh}{n \vw(\rs)^2}\left(\frac{9-\Gamma}{2} \frac{\Mstar(\rs)}{\Mbh} +\frac{7}{2}\right),
\label{eq:stag_analytic})
\end{align}
in terms of the radius of the sphere of influence (eq.~[\ref{eq:rsoi}]), where $n$ is the power-law slope of the density profile at $r = r_{\rm s}$.

If the SMBH dominates the gravitational potential at the stagnation radius ($M_{\bullet} \gg M_{\star}$; $v_{w,0} \gg \sigma$) then equation (\ref{eq:stag_analytic}) simplifies greatly
\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n v_{\rm w}^{2}} \approx 6
 \, \pc \,\, \Mbheight v_{500}^{-2},
  \label{eq:stag_simple}
\end{align}
where the numerical estimate assumes $n = 1$.  This expression is similar to that obtained by \citet{Volonteri+11} on heuristic grounds (their equation~[6]).  In the opposite limit when the wind heating is weaker ($v_{\rm w,0} \ll \sigma$), the stagnation radius moves to very large radii or may not exist at all.

\subsubsection{SMBH Accretion Rate}

The SMBH accretion rate is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have made use of equation (\ref{eq:stag_simple}), i.e. assuming that $v_{w,0} \gg \sigma$.  The resulting Eddington ratio $\eddr$ is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core} \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
To reproduce the estimated values $\sim 10^{-XX}$ in low luminosity AGN, we thus require $v_{w,0} \sim $.  {\bf BDM: put some realistic numbers, e.g. from Ho 2009 or better references}

>>>>

\begin{abstract}
We calculate steady-state one-dimensional hydrodynamic profiles of hot gas in the nuclei of quiescent galaxies for a range of black hole masses, parameterized gas heating rates, and physically-motivated stellar density profiles.  Mass is supplied to the circumnuclear medium (CNM) by stellar winds, while gas is heated by a combination of stellar winds, supernovae, and black hole feedback.  We derive analytic estimates for quantities including the stagnation radius where the radial velocity goes to zero (if it exists) and the black hole accretion rate as a function of the heating efficiency, the latter of which is related to the star-formation history of the stellar population.  These results (neglecting radiative cooling) are used to assess over what range of conditions radiative instability will develop, assuming an average star formation history for galaxies of the given black hole mass.  Our main conclusion is that.....  We briefly discuss implications of our results for estimates of the accretion rate using Bondi formula, and for the diversity of CNM environments encountered by the jetted outflows from tidal disruption events, and for the use of an assumed $\dot{M}-M_{\bullet}$ relationship to constrain the occupation fraction of SMBHs in low mass galaxies.
\end{abstract}

@@

To determine why most SMBHs appear to be inactive requires an understanding of their gaseous environments.  Hot gas near the sphere of influence of the SMBH (hereafter the `circumnuclear medium', or CNM) controls the mass accretion rate, $\dot{M}$.  The latter in turn sets the luminosity of the accretion flow and the impact of its energy and momentum input (`feedback') on larger physical scales.  The presence of dense gas in the nucleus may also lead to runaway cooling, resulting in bursty star-formation and AGN activity (reference Ciotti and others).

Knowledge of how $\dot{M}$ varies with the SMBH mass, $\Mbh$, and other properties of the
galactic nucleus informs key questions, such as the co-evolution of SMBHs and their host galaxies with cosmic time.  SMBH growth in the low redshift universe appears to be dominated by lower mass black holes (e.g.~\citealt{Heckman+04}).  This is often
attributed to cosmological `down sizing' (e.g, \citealt{Gallo+08}), even though the physical processes involved may be distinct from those operating at higher SMBH masses and accretion rates in quasars.

The accretion rates of low mass SMBHs, and their resulting X-ray luminosities $L_X$, also have implications for what observational constraints can be placed on their occupation fraction in low mass galaxies and the possible existence of intermediate mass black holes.  By combining the average relationship between $\Mbh$ and
$L_X$ measured from a sample of high mass galaxies (\citealt{Gallo+08}) with upper limits on $L_X$ in lower mass
galaxies, \citet{MillerGallo+:2014a}  infer tentative evidence for an occupation fraction less than unity in the latter.  This method, however, relies on extrapolating a power-law fit to the $L_X-\Mbh$ distribution to low $\Mbh$ below the detection threshold.

@@

Past studies, employing multi-dimensional numerical hydrodynamics and including variety of (parameterized) physical effects, have focused on massive elliptical galaxies (Ciotti refs).  A key finding of these works is the periodic development of cooling instabilities, which temporarily increase the gas accretion rate until black hole feedback becomes sufficiently important to shut off the flow and halt SMBH growth.  Here we instead focus on time-independent models, neglecting radiative cooling.  This approach allows us to systematically explore the relevant parameter space and to derive simple analytic expressions that prove useful in exploring whether cooling instabilties manifest across the expected range of galaxy properties, or whether other (non-AGN) forms of feedback can result in steady accretion.  Even if cooling instabilties develop on galactic scales over long timescales, quasi-steady accretion may persist for briefer periods on small scales.  We are thus motivated to explore whether steady flow is achieved under any circumstances, or whether accretion is always building towards instability.  
%A chief goal of our analysis is to provide a framework for interpreting the properties of quiescent galactic nuclei given the comparatively well understood physics of stellar winds and supernova heating.

In the presence of strong heating, one-dimensional steady-state flow is characterized by an inflow-outflow structure, with a critical radius known as the``stagnation radius" $\rs$ where the radial velocity passes through zero.  Mass loss from stars interior to the stagnation radius is accreted, while that outside $\rs$ is unbound in an outflow from the nucleus.  The stagnation radius, rather than the Bondi radius, thus controls the SMBH accretion rate (although we will show that $\rs$ can in many cases lie within a factor of few of the nominal Bondi radius).  In the case of sufficiently weak heating, however, no stagnation radius may exist at all, i.e. the entire ISM of the galaxy is feeding the SMBH.  Another goal of this work is to assess under what conditions a stagnation radius exists, as this question is intimately tied to that of cooling instability.  
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion.

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to measure the radial surface brightness profiles for hundreds of nearby early type galaxies. The measured brightness profile is generally well-fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.  Assuming the stellar population is spherically symmetric, this corresponds to a stellar density $\rhostar \propto r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for $r \gg \rb$.  To good approximation
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star(\rsoi) \left(r/\rsoi\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star(r_b) \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star(\rsoi)$ is the stellar density at the radius of the black hole sphere of influence (soi), 
\be
\rsoi \simeq G \Mbh/\sigma^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\ee
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$ relationship of \citet{Gultekin+09},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
The use of equation (\ref{eq:Msigma}) is questionable for low mass black holes (e.g., \citealt{Greene&Ho07}), but our results are not sensitive to this assumption.

A galaxy model is fully specified by four parameters: $\Mbh$, $\Gamma$, $r_b$,
and $\beta$.  We compute models for three different black hole masses, $\Mbh = 10^6$, $10^7$, $10^8 \Msun$.  The distribution of $\Gamma$ in the \citet{LauerFaber+:2007a} sample is bimodal, with a concentration of ``core" galaxies with $\Gamma \lsim 0.3$ and a concentration of ``cusp" galaxies with $\Gamma \gsim 0.7$.  We bracket these possibilities by considering models with $\Gamma=0.8$ and $\Gamma=0.1$.  Since our results are most sensitive to the properties of the gas flow on radial scales $\ll r_b$, we fix $\beta = 2$ (a typical value).  However, the presence of $\rb$ is necessary for to obtain a converged steady state for some regions of our parameter space. We adopt $\rb=100$ pc for cusp galaxies ($\Gamma=0.8$), while for core galaxy models we employ the scaling relation $\rb = $, motivated again by the \citet{LauerFaber+:2007a} sample. ({\bf BDM: fill in!}).  Though simplistic, our results are not sensitive to our assumption about the value of $\rb$.

@@

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the density $\rho$, temperature $T$, and radial velocity $v$ of the CNM for each galaxy model by solving the equations of one-dimensional, time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial
      v}{\partial r}\right) =-\frac{\partial p}{\partial r}- \rho\frac{GM_{\rm enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] 
, 
\label{eq:dsdt}
\end{align}
where $p$ and $s$ are the pressure and specific entropy, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p = \rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$.

The source term in the continuity equation (eq.~[\ref{eq:drhodt}]),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years.  To good approximation $\eta\simeq 0.02 (t_{\star}/t_h)^{-1.3}$ at time $t_{\star}$ following a star burst (e.g., \citealt{Ciotti+91}),\footnote{\citet{Ciotti+91} give the overall mass return rate from evolved stars as a function of B-band luminosity.  Although we give the mass return rate per unit volume, our expressions are equivalent.} although $\eta$ can be higher if star formation is ongoing (Fig X).

Source terms $\propto q$ also appear in the momentum and entropy
equations (eqs.~[\ref{eq:dvdt}] and [\ref{eq:dsdt}]) because isotropic
mass injection in the SMBH rest frame represents a source of momentum
and energy relative to the mean flow.  Physically, these result from
the mismatch between the properties of virialized gas injected locally
by stellar winds and the mean properties of the background flow.  The
term $\propto -q c_{s}^{2}$ acts as an important agent in stabilizing the flow to runaway cooling ($\S\ref{sec:instability}$).

@@

To isolate the physics of interest, our baseline calculations neglect two potentially important effects: heat conduction and radiative cooling.  Heat transport by electron conductivity is included with an additional term to the right hand side of the entropy equation of the form  
\be \dot{q}_{\rm cond} = \nabla\cdot(\kappa \nabla T), \ee
where $\kappa_{\rm
  eff}=\kappa_{\rm spitz}/(1+\psi)$ (\citealt{DaltonBalbus:1993a}) is
the conducitivity and $\kappa_{\rm spitz} = \kappa_0 T^{5/2}$ is the
classical \citet{Spitzer62} value ($\kappa_0\simeq 2\times 10^{-6}$ in
cgs units).  The flux limitor $\psi = \kappa_{\rm spitz} \nabla
T/(5\phi \rho c_s^3)$ saturates the conductive flux if the mean free
path for electron coulomb scattering becomes comparable to the
temperature length scale, where $c_s \equiv (kT/\mu m_p)^{1/2}$ is the
isothermal sound speed and $\phi \lesssim 1$ is an uncertain
dimensionless constant (we adopt $\phi = 0.1$).  Even a weak magnetic
field perpendicular to the flow could in principle suppress the
conductivity.  However, the radially-decreasing temperature profiles
of interest are susceptible to the magneto-thermal instability
(\citealt{Balbus01}), the non-linear evolution of which is believed to
result in a radially-directed field structure
(\citealt{Parrish&Stone07}).

Radiative cooling is also not included in our baseline calculations.  Across much of
the parameter space radiative cooling is subdominant compared to other
sources of cooling.  Radiative cooling cannot be neglected when $\vwO$
is low (and/or $\eta$ is high), but once important it leads to thermal
instabilities (e.g.~\citealt{McCourt+12}) which our 1D
time-independent model cannot account for self-consistently anyways.
Indeed, one of the main goals of this work is to use our analytic and
numerical solutions neglecting radiation to ascertain over what
physical conditions cooling instabilities are likely to develop
($\S\ref{sec:cooling}$).

@@

@@

\section{Analytic Results}
\label{sec:results}

We begin by describing analytic estimates of physical quantities, such as the stagnation radius $\rs$ and the mass accretion rate, the derivation of which are given in Appendix \ref{app:rs}.  Our results are expressed in terms of the SMBH mass $M_{\bullet} = 10^{8}M_{\bullet,8}M_{\odot}$ and the wind heating parameter
(eq.~[\ref{eq:dsdt}])
\begin{eqnarray}
  v_{\rm w} &=& v_{\rm w,0}\left(1+\frac{\sigma^{2}}{v_{\rm w,0}^{2}}\right)^{1/2}  \nonumber \\
  &\approx& v_{\rm w,0}\left(1+0.14 \Mbheight^{0.43}\left[\frac{v_{w,0}}{500\,\rm km\,s^{-1}}\right]^{-1}\right)^{1/2},
\end{eqnarray}
where the final equality makes use of $M_{\bullet}-\sigma$ (eq.~[\ref{eq:Msigma}]).

\subsection{Flow Properties Near the Stagnation Radius}

Continuity of the entropy derivative at the stagnation radius demands that the temperature at the stagnation radius is given by (eq.~[\ref{eq:first_law}])
\begin{align}
T \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p v_{w}^{2}}{2} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
where $v_{500} \equiv v_{\rm w}/500$ km s$^{-1}$.

Hydrostatic equilibrium can be used to determine the location of the stagnation radius where $v = 0$ (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
\rs=\frac{G \Mbh}{n \vw(\rs)^2}\left(\frac{9-\Gamma}{2} \frac{\Mstar(\rs)}{\Mbh} +\frac{7}{2}\right),
\label{eq:stag_analytic}
\end{align}
where $n$ is the power-law slope of the density profile at $r = r_{\rm s}$.  If the SMBH dominates the gravitational potential at the stagnation radius then $v_{w,0} \gg \sigma$ and $r_{\rm s} < r_{\rm soi}$ (eq.~[\ref{eq:rsoi}]).  In this case equation (\ref{eq:stag_analytic}) simplifies to
\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n v_{\rm w}^{2}} \approx 6
 \, \pc \,\, \Mbheight v_{500}^{-2},
  \label{eq:stag_simple}
\end{align}
where the numerical estimate assumes $n = 1$.  This expression is similar to that obtained by \citet{Volonteri+11} on heuristic grounds (their eq.~6).  In the opposite limit of weak heating ($v_{\rm w,0} \ll \sigma$), the stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not exist at all.  This makes intuitive sense: gas is injected by stars which are gravitationally bound to the black hole, so outflows only occur if the heating rate per gram $\sim v_{\rm w}^{2}$ exceeds the specific gravitational binding energy $\sim \sigma^{2}$.

\subsection{SMBH Accretion Rate}

The SMBH accretion rate is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have made use of equation (\ref{eq:stag_simple}), assuming that $v_{w,0} \gg \sigma$.  The resulting Eddington ratio $\eddr$ is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core} \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w,0} \gtrsim 150M_{\bullet,8}^{0.5}$
